|

### 2022-12-28 drafts/fa_2022/dcr_white_identity_relative_deprivation_fa_2022.qmd
fix languages

```{r}
#| label: load-data-block
# create list object for datasets
data = list()
# Create empty lasso list to store fitted models
lasso = list()
# Specify number of iterations for imputation
data_sets = 10
# load the cleaned data
    #* 2012
data[['anes_2012']] <- read.csv('../../data/anes-2012/anes-2012-updated.csv') |>
    select(wid, rboff, epast, unpast, edu, income, losejob, immtakejobs, ecfamily, ecjob, tradbreak, govbias, inflwhite, whitediscrim, raceResent, pid, female)
    #* 2016
data[['anes_2016']] <- read.csv('../../data/anes-2016/anes-2016-updated.csv') |>
    select(wid, rboff, epast, unpast, edu, income, losejob, immtakejobs, ecfamily, ecjob, tradbreak, govbias, inflwhite, whitediscrim, raceResent, pid, female)
    #* 2020
data[['anes_2020']] <- read.csv('../../data/anes-2020/anes-2020-updated.csv') |>
    select(wid, rboff, epast, unpast, edu, income, losejob, immtakejobs, ecfamily, ecjob, govbias, inflwhite, whitediscrim, raceResent, pid, female)
```

@@

In many eras of American history, there are several instances where we have sought to reclassify who is white. During the antebellum period, whiteness was no longer simply one's economic class. However, now it had become a reflection of a racial hierarchy that afforded some various rights and excluded others from enjoying the same rights (e.g., literacy tests). During this era, rather than conceptualizing whites as landowners of Caucasian descent, the definition of what was white shifted to include those previously labeled non-white Europeans, such as the Irish. Changes to definitions of who is white illustrate that the primary motivator was race and not class as to who holds some rights and who does not [@rile-hayward_2013].

More contemporaneously, whites hold several strong political attitudes racialized by politicians. Whites' attitudes toward welfare programs are not too distinct from those toward Black Americans [@gilens_1999]. A provided explanation is that social welfare policies are more unpopular among whites when the media frames these policies as benefiting non-whites [@winter_2008;@van-doorn_2015]. Explicit cues reminding individuals of the racial hierarchy entrench the recipients' attitudes toward Black Americans [@lyle_2014; see also @wellman_et-al_2016]. Conservatives argued that Barack Obama, the first Black president, expressed support for the Affordable Care Act as a way to disproportionately help Black Americans [@tesler_2012;@tesler_2016]. These messages corresponded with a rise of survey reports expressing old-fashioned racism among whites during his administration [@tesler_2013]. While whites tend to report negative views about government aid for those in need when they are non-white recipients, evidence suggests that whites who live in communities with a higher proportion of poor whites report much more positive views of the poor than those that perceive the poor as mostly non-white [@hopkins_2009]. Additionally, telling whites that other whites will be the beneficiaries of an entitlement program increases their support for such a program [@wetts_willer_2018].

@@

Not all agree, however. Some suggest that the negative attitudes that whites express toward immigrants result from concerns about cultural changes in those living in border towns [@hainmueller_hopkins_2014]. Others suggest that simply reminding whites of these demographic shifts activates these racial identity-based attitudes [@craig_richeson_2014]. Some other evidence suggests that these identity-based attitudes are better predicted by ethnic antagonism [@bartels_2020] and by racial animus [@fahey_et-al_2020] and explicitly take the position that a reaction to economic threats does not drive these. However, these attitudes tend to focus on prejudiced views rather than a reflection of a connection to other whites.

# Is white identity non-material?

@@

As this is an open question and is rather a complex one, I attempt to wade into this by parsing through which of these two perspectives are more plausible. As I discuss in the concluding section, I argue that scholars should continue taking the conceptualization and measurement of white political identification seriously and examine the mechanisms that connect economic and political threats to white political identity. In the next section, I examine whether attitudes expressing concern with economic or political threats are more robust correlates of the degree to which someone feels that being white is important to their identity. The goal is not to test particular causal mechanisms but to examine whether my belief and others' that white identity reflects reactions to political threats more than it does economic.

@@

data[['anes_2012']] |>
  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Immigrants take jobs` = immtakejobs, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `New lifestyles breaking society` = tradbreak, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
  datasummary_skim(notes = c('NAs are included in Unique column.', 'Data source: 2012 American National Election Study, unweighted.'))
```

@@

data[['anes_2016']] |>
  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Immigrants take jobs` = immtakejobs, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `New lifestyles breaking society` = tradbreak, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
  datasummary_skim(notes = c('NAs are included in Unique column.', 'Data source: 2016 American National Election Study, unweighted.'))
```

@@

data[['anes_2020']] |>
  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Immigrants take jobs` = immtakejobs, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
  datasummary_skim(notes = c('NAs are included in Unique column.', 'Data source: 2020 American National Election Study, unweighted.'))
```

@@

rows <- tribble(~term, ~`2012 LWD`, 'Economic', '', 'Non-material', '', 'Demographics', '', 'Thresholds', '')
attr(rows, 'position') <- c(1, 10, 19, 28)

@@

```{r}
#| label: tbl-2016-lasso-simple
#| tbl-cap: Predictors of white identity in 2012

rows <- tribble(~term, ~`2016 LWD`, 'Economic', '', 'Non-material', '', 'Demographics', '', 'Thresholds', '')
attr(rows, 'position') <- c(1, 13, 22, 31)

@@

rows <- tribble(~term, ~`2020 LWD`, 'Economic', '', 'Non-material', '', 'Demographics', '', 'Thresholds', '')
attr(rows, 'position') <- c(1, 13, 22, 31)

@@

model <- list(
    '2012 ANES' = lasso[['2012']],
    '2016 ANES' = lasso[['2016']],
    '2020 ANES' = lasso[['2020']]
)
rows <- tribble(~term, ~`2012 ANES`, ~`2016 ANES`, ~`2020 ANES`, 'Economic', '', '', '','Non-material', '', '', '', 'Demographics', '', '', '', 'Thresholds', '','', '')
attr(rows, 'position') <- c(1, 13, 22, 31)

>>>>

```{r}
#| label: load-data-block
# create list object for datasets
data = list()
# Create empty lasso list to store fitted models
lasso = list()
# Specify number of iterations for imputation
data_sets = 10
# load the cleaned data
    #* 2012
data[['anes_2012']] <- read.csv('../../data/anes-2012/anes-2012-updated.csv') |>
    select(wid, rboff, epast, unpast, edu, income, losejob, ecfamily, ecjob, tradbreak, govbias, inflwhite, whitediscrim, raceResent, pid, female)
    #* 2016
data[['anes_2016']] <- read.csv('../../data/anes-2016/anes-2016-updated.csv') |>
    select(wid, rboff, epast, unpast, edu, income, losejob, ecfamily, ecjob, tradbreak, govbias, inflwhite, whitediscrim, raceResent, pid, female)
    #* 2020
data[['anes_2020']] <- read.csv('../../data/anes-2020/anes-2020-updated.csv') |>
    select(wid, rboff, epast, unpast, edu, income, losejob, ecfamily, ecjob, govbias, inflwhite, whitediscrim, raceResent, pid, female)
```

@@

@@

@@

As this is an open question and is rather a complex one, I attempt to wade into this by parsing through which of these two perspectives are more plausible. In the next section, I examine whether attitudes expressing concern with economic or political threats are more robust correlates of the degree to which someone feels that being white is important to their identity. The goal is not to test particular causal mechanisms but to examine whether my belief and others' that white identity reflects reactions to political threats more than it does economic.

@@

data[['anes_2012']] |>
  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `New lifestyles breaking society` = tradbreak, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
  datasummary_skim(notes = c('NAs are included in Unique column.', 'Data source: 2012 American National Election Study, unweighted.'))
```

@@

data[['anes_2016']] |>
  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `New lifestyles breaking society` = tradbreak, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
  datasummary_skim(notes = c('NAs are included in Unique column.', 'Data source: 2016 American National Election Study, unweighted.'))
```

@@

data[['anes_2020']] |>
  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
  datasummary_skim(notes = c('NAs are included in Unique column.', 'Data source: 2020 American National Election Study, unweighted.'))
```

@@

rows <- tribble(~term, ~`2012 LWD`, 'Economic', '', 'Non-material', '', 'Demographics', '', 'Thresholds', '')
attr(rows, 'position') <- c(1, 18, 27, 32)

@@

```{r}
#| label: tbl-2016-lasso-simple
#| tbl-cap: Predictors of white identity in 2016

rows <- tribble(~term, ~`2016 LWD`, 'Economic', '', 'Non-material', '', 'Demographics', '', 'Thresholds', '')
attr(rows, 'position') <- c(1, 18, 27, 32)

@@

rows <- tribble(~term, ~`2020 LWD`, 'Economic', '', 'Non-material', '', 'Demographics', '', 'Thresholds', '')
attr(rows, 'position') <- c(1, 18, 25, 30)

@@

model <- list(
    '2012 ANES' = lasso[['2012']],
    '2016 ANES' = lasso[['2016']],
    '2020 ANES' = lasso[['2020']]
)
rows <- tribble(~term, ~`2012 ANES`, ~`2016 ANES`, ~`2020 ANES`, 'Economic', '', '', '','Non-material', '', '', '', 'Demographics', '', '', '', 'Thresholds', '','', '')
attr(rows, 'position') <- c(1, 18, 27, 32)
